CYO Project Report - Brazilian Ecommerce

### 2021-07-13 finalReport.Rmd
tested model with customer loc effect aand there is improvement in the model knn, gam and rf

# so raw data
raw_data <- orders_dataset %>% 
  inner_join(order_items_dataset,by='order_id') %>%
  inner_join(sellers_dataset,by='seller_id') %>% 
  inner_join(customers_dataset,by='customer_id') %>%  
  inner_join(products_dataset,by='product_id') %>%
  mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),
         product_id=as.factor(product_id), 
         seller_id=as.factor(seller_id),
         seller_zip_code_prefix=as.factor(seller_zip_code_prefix),
         customer_unique_id=as.factor(customer_unique_id),
         customer_zip_code_prefix=as.factor(customer_zip_code_prefix),
         product_category_name=as.factor(product_category_name)) %>%
  select(-customer_id,-order_status,-seller_city,-customer_city, -payment_approved_at,-order_id,-order_item_id,-product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
#  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-customer_state, -payment_approved_at,-order_id, -order_item_id, -product_id,-seller_zip_code_prefix,-customer_unique_id,-customer_zip_code_prefix)
```

@@

```{r Cleanup and save required dataset, include=FALSE, echo=FALSE}
save(wrangled_data,file='rda/wrangled_data.rda')
rm(geolocation_dataset, order_items_dataset, orders_dataset, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore, raw_data)
```

@@

@@

# ensure validation_dataset has products only for concerned product categories and seller_ids which are in training_dataset 
validation_dataset <- validation_dataset %>%
  semi_join(training_dataset,by='product_category_name') %>%
  semi_join(training_dataset,by='seller_id')
# validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')

@@

We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
```{r Split data for training, include=FALSE}
# Let's split training_dataset into two parts
set.seed(1,sample.kind = 'Rounding')
training_dataset <- training_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer()

@@

train_subset <- training_dataset[-test_indices,]
test_subset <- training_dataset[test_indices,]
# ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id')
#test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')

@@

@@

  train_subset_for_encoding <- train_subset[1:30000,]
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

@@

```{r Apply Target encoding for customer_state, echo = FALSE}
  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-seller_state)
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

@@

```{r Apply Target encoding for seller_state, echo = FALSE}
  train_subset_for_encoding <- train_subset[1:40000,] 
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

@@

@@

  test_weight_set <- test_weight_set %>%
    #semi_join(train_weight_set,by='product_category_name') %>%
    semi_join(train_weight_set,by='seller_state')

  weight_rmses <- sapply(weights, function(weight) {
      seller_state_effects <- train_weight_set %>% 
        group_by(seller_state) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n) %>% rename()

      train_weight_set <- train_weight_set %>% left_join(seller_state_effects,by='seller_state') %>% rename(seller_state_effect=mean) %>% select(-seller_state) #-product_category_name

      test_weight_set <- test_weight_set %>% left_join(seller_state_effects,by='seller_state') %>% rename(seller_state_effect=mean) %>% select(-seller_state) # -product_category_name,

@@

weight_to_use <- mean(weight_to_use_per_partition)  
# 15
```

```{r Post regularization with seller state effect, echo=FALSE}

@@

seller_state_effects_list <- lapply(1:10, function(index) {
    train_weight_set <- train_subset[-test_weight_indices[,index],]

    seller_state_effects <- train_weight_set %>% 
      group_by(seller_state) %>%
      summarise(mean=mean(number_of_hours),n=n()) %>%
      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
      select(-n) %>% rename()

    seller_state_effects
  }
)

sse <- seller_state_effects_list[[1]]
for (index in 2:10) {
   sse <- sse %>% full_join(seller_state_effects_list[[index]],by='seller_state', suffix=c("1",index))
}

# whereever value is NA, we have used global mean
sse <- sse %>% mutate(across(-seller_state,~if_else(is.na(.x),global_mean,.x)))

# as different partitions may not have exactly same data set combined the results
seller_state_effects <- tibble(sse[,1], seller_state_effect=rowMeans(sse[,-1],na.rm = TRUE))

train_subset <- train_subset %>% left_join(seller_state_effects,by='seller_state') %>% select(-seller_state) %>% mutate(seller_state_effect=if_else(is.na(seller_state_effect),global_mean,seller_state_effect))

test_subset <- test_subset %>% left_join(seller_state_effects,by='seller_state') %>% select(-seller_state)
```

@@

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
#190.243 for k=41. The seller_state is not of much help. it decrease RMSE by .14

@@

@@

```{r rf model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 2, p = .8)
rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250)
#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200

@@

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 181.8166 for mtry 15. Technically there are total of 82 variables due to 73 product categories and 9 other predictors. The seller_state is not of much help. it increases RMSE by .6. So we should consider ignoring it.

@@

```{r Gam model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number =2, p = .8)
gam_fit <- train(number_of_hours~., train_subset, method='gam',trControl=tuningControl) # check for tuning parameters

@@

predictions <- predict(gam_fit,test_subset %>% select(-number_of_hours))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 195.3235

>>>>

# so raw data
raw_data <- orders_dataset %>% 
  inner_join(order_items_dataset,by='order_id') %>%
  inner_join(sellers_dataset,by='seller_id') %>% 
  inner_join(customers_dataset,by='customer_id') %>%  
  inner_join(products_dataset,by='product_id') %>%
  mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),
         product_id=as.factor(product_id), 
         seller_id=as.factor(seller_id),
         seller_zip_code_prefix=as.factor(seller_zip_code_prefix),
         customer_unique_id=as.factor(customer_unique_id),
         customer_zip_code_prefix=as.factor(customer_zip_code_prefix),
         product_category_name=as.factor(product_category_name)) %>%
  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city, -payment_approved_at,-order_id,-order_item_id,-product_id,-seller_zip_code_prefix,-customer_unique_id)
```

@@

```{r Cleanup and save required dataset, include=FALSE, echo=FALSE}
save(wrangled_data,file='rda/wrangled_data.rda')
rm(geolocation_dataset, order_items_dataset, orders_dataset, product_category_name_translation, products_dataset, sellers_dataset, sellers_to_ignore, customers_dataset, customers_to_ignore)
# TODO delete raw_data as well.
```

@@

```{r Method to reduce levels in customer_zip_code_prefix, include = FALSE}
# trimming zip code to three digits to be able to extract zip code component which represents region, sub-region and sector data.
compute_rough_customer_location <- function(df) {
  df %>% mutate(customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(3),pad=c('0'),side='right')) %>% 
    mutate(customer_loc = as.numeric(str_extract(customer_zip_code_prefix,"\\d\\d\\d"))) 
  #%>% select(-customer_zip_code_prefix)
}
```

@@

# ensure validation_dataset has products only for concerned product categories and seller_ids which are in training_dataset 
validation_dataset <- validation_dataset %>%
  semi_join(training_dataset,by='product_category_name') %>%
  semi_join(training_dataset,by='seller_id') %>% 
  semi_join(training_dataset,by="customer_zip_code_prefix")
## TODO review if we want to change this join
# validation_dataset <- validation_dataset %>% semi_join(training_dataset,by='seller_id')

@@

We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
```{r Split data for training, include=FALSE}
# Let's split training_dataset into two parts
set.seed(1,sample.kind = 'Rounding')
training_dataset <- training_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer() %>% compute_rough_customer_location()

@@

train_subset <- training_dataset[-test_indices,]
test_subset <- training_dataset[test_indices,]
# ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>% semi_join(train_subset,by='seller_id') %>% 
  semi_join(train_subset,by="customer_zip_code_prefix")
#test_subset <- test_subset %>% semi_join(train_subset,by='seller_id')

@@

## Explore truncated customer_zip_code_prefix with number_of_hours

```{r Explore truncated customer_zip_code_prefix with number_of_hours}
# let's plot errorbar for number of hours vs truncated customer_zip_code_prefix
training_dataset %>% mutate(czc = as.numeric(str_extract(customer_zip_code_prefix,"\\d\\d\\d"))) %>% group_by(czc) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(czc=reorder(czc,mean)) %>% ggplot(aes(x=czc,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```

@@

  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-customer_loc, -customer_zip_code_prefix, -customer_state)

  global_mean <- mean(train_subset_for_encoding$number_of_hours)

@@

```{r Apply Target encoding for customer_state, echo = FALSE}
  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc,-customer_zip_code_prefix)
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

@@

```{r Apply Target encoding for customer_loc, echo = FALSE}
  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_zip_code_prefix)
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

@@

  weights <- seq(0,10,5)

@@

  test_weight_set <- test_weight_set %>%
    #semi_join(train_weight_set,by='product_category_name') %>%
    semi_join(train_weight_set,by='customer_loc')

  weight_rmses <- sapply(weights, function(weight) {
      customer_loc_effects <- train_weight_set %>% 
        group_by(customer_loc) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n) %>% rename()

      train_weight_set <- train_weight_set %>% left_join(customer_loc_effects,by='customer_loc') %>% rename(customer_loc_effect=mean) %>% select(-customer_loc) #-product_category_name

      test_weight_set <- test_weight_set %>% left_join(customer_loc_effects,by='customer_loc') %>% rename(customer_loc_effect=mean) %>% select(-customer_loc) # -product_category_name,

@@

weight_to_use <- mean(weight_to_use_per_partition) 
# 5/3 = 1.66667
```

```{r Post regularization with customer loc effect, echo=FALSE}

# prepare customer_loc_effect based on regularisation weight we have identified
  global_mean <- mean(train_subset$number_of_hours)

@@

customer_loc_effects_list <- lapply(1:10, function(index) {
    train_weight_set <- train_subset[-test_weight_indices[,index],]

    customer_loc_effects <- train_weight_set %>% 
      group_by(customer_loc) %>%
      summarise(mean=mean(number_of_hours),n=n()) %>%
      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
      select(-n) %>% rename()

    customer_loc_effects
  }
)

cle <- customer_loc_effects_list[[1]]
for (index in 2:10) {
   cle <- cle %>% full_join(customer_loc_effects_list[[index]],by='customer_loc', suffix=c("1",index))
}

# whereever value is NA, we have used global mean
cle <- cle %>% mutate(across(-customer_loc,~if_else(is.na(.x),global_mean,.x)))

# as different partitions may not have exactly same data set combined the results
customer_loc_effects <- tibble(cle[,1], customer_loc_effect=rowMeans(cle[,-1],na.rm = TRUE))

train_subset <- train_subset %>% left_join(customer_loc_effects,by='customer_loc') %>% select(-customer_loc,-customer_zip_code_prefix) %>% mutate(customer_loc_effect=if_else(is.na(customer_loc_effect),global_mean,customer_loc_effect))

test_subset <- test_subset %>% left_join(customer_loc_effects,by='customer_loc') %>% select(-customer_loc,-customer_zip_code_prefix)
```

@@

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
#184.7906 for k=34.

@@

```{r correlation, eval=FALSE}
cor(train_subset %>% select(price,freight_value,customer_loc_effect,customer_state_effect,seller_id_effect,distance_between_two_points,mean_delay_in_handover,volume,number_of_hours,product_weight_g),use="pairwise.complete.obs") %>% knitr::kable()
```

@@

```{r rf model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 2, p = .8)
rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 40, 10)),trControl=tuningControl,ntree=250)
#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200

@@

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 175.0988 for mtry 15. Technically there are total of 82 variables due to 73 product categories and 9 other predictors. The seller_state was not of much help. it was actually increasing RMSE by .6. So have ignored it.

@@

```{r Gam model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')
preProcessScalingValues <- preProcess(train_subset %>% select(-number_of_hours), method = c("center", "scale"))

tuningControl=trainControl(method = "cv", number =3, p = .8)
gam_fit <- train(number_of_hours~., scaled_train_subset, method='gam',trControl=tuningControl) # check for tuning parameters

@@

predictions <- predict(gam_fit,scaled_test_subset %>% select(-number_of_hours))

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
# 185.1074 uses GCV.Cp method and select=FALSE
